\documentclass{article}

\begin{document}

Crystal Mandal

\subsection*{Acts and Revivals:}
\begin{itemize}
  \item \textbf{Revival of the Ku Klux Act}
  \item \textbf{The Civil Rights Act of 1964}
  \item \textbf{The Voting Act of 1965}
\end{itemize}

\subsection*{Investigate the Act:}

  \begin{itemize}
    \item \textit{Ku Klux Klan Revival}: The Ku Klux Klan Revival (in the mid 20th
    century as a ``third" Klan) was a response to the Growing Civil Rights Movement. 
    The Ku Klux Klan aimed not to protect but to spread White Supremacy and White 
    Nationalist ideals in America.
    \item \textit{The Civil Rights Act of 1964}: The Civil Rights Act of 1964 was 
    originally proposed in response to a wave of protests in 1963. It establishes 
    Federal Protections against discrimination based on race, color, religion, sex, 
    and national origin. 
    \item \textit{The Voting Act of 1965}: The Voting Act of 1965 prohibits racial 
    discrimination in voting. It enforces the voting rights protected by the Fourteenth 
    and Fifteenth Ammendment.  
  \end{itemize}

\subsection*{Evaluate Impact:}

  \begin{itemize}
    \item \textit{Ku Klux Klan Revival}: I'd rather not dwell on how successful the 
    Klan is in spreading their ideology, but it is quite difficult not to notice the 
    rising White Nationalist trends in contemporary American Politics.
    \item \textit{The Civil Rights Act of 1964}: The Civil Rights Act of 1964 is 
    widely regarded as one of the most important legislative achievements in American 
    History. Though some, I included, would argue that the ``spirit" of the Civil 
    Rights Act is yet to be properly established into American Law, it is undecidedly 
    foundational legislature for creating equity between all American Peoples.
    \item \textit{The Voting Act of 1965}: The US Department of Justice proudly states 
    that The Voting Act of 1965 is the single most effective piece of civil rights 
    legislation passed by Congress.
  \end{itemize}

\subsection*{Historical Significance:}

  \begin{itemize}
    \item \textit{Ku Klux Klan Revival}: The Klan has remained an important historical 
    measure of Anti-Black Sentiment in America, particularly in the South. In 1963, 
    members of the ``Third Klan" bombed the 16th Street Baptist Church in Birmingham, 
    Alabama, which led directly to more protests and, eventually, the Civil Rights Act 
    of 1964.
    \item \textit{The Civil Rights Act of 1964}: Happily, the Civil Rights Act has not 
    been repealed, nor has repeal been threatened. Unfortunately, the monumental DEI 
    policies which follow in the ``spirit" of the Civil Rights Act are now under attack.
    \item \textit{The Voting Act of 1965}: The Voting Act of 1965 Secured voting rights 
    for all eligible citizens whilst simultaneously prohibiting any racial bias in 
    the eligibility constraints for voting. Thankfully, this act has, to my knowledge, 
    not been criticized recently. 
  \end{itemize}

\end{document}
